---
title: "Data Exploration Project"
author: "George T"
date: "2/17/2022"
output: html_document
---
# loading packages
```{r}
library(tidyverse)
library(purrr)
library(car)
library(fmtr)
```


# reading in data files
```{r}
setwd("~/Downloads/Lab3_Rawdata")

trends <- list.files(pattern = "trends_", full.names = TRUE) %>% 
  lapply(read_csv) %>% 
  bind_rows() 
  
scorecard <- read.csv("Most+Recent+Cohorts+(Scorecard+Elements).csv", header = TRUE)

id_name_link <- read.csv("id_name_link.csv", header = TRUE)
```

```{r}
id_name_link <- id_name_link%>%
  group_by(schname)%>%
  mutate(N = n())%>%
  filter(N == 1)
```


# linking together the data
```{r}
trends_and_id_name <- left_join(x = id_name_link, y = trends, by = "schname")

ourdata <- left_join(x = trends_and_id_name, y = scorecard, by = c("unitid" = "UNITID"))
```


# cleaning

# remove schools with the same name
```{r}
ourdata_norepeat <- ourdata %>%
  group_by(unitid) %>%
  slice(1)
```

# filtering out colleges that don't predominantly grant bachelors degrees
```{r}
ourdata_norepeat <- ourdata_norepeat %>%
  filter(PREDDEG == 3)
```


# Analysis


# grouping by monthorweek
```{r}
month_or_week <- ourdata %>%
  group_by(monthorweek) %>%
  summarize(n = n())
```

# remove dates with NA
```{r}
month_or_week_clean <- na.omit(month_or_week, c("monthorweek"))
```

# dates before sept 15th
```{r}
month_or_week_clean$monthorweek
before_sept <- month_or_week_clean$monthorweek[1:128]
before_sept <- as.data.frame(before_sept)
```
# dates after sept 2015
```{r}
month_or_week_clean$monthorweek
after_sept <- month_or_week_clean$monthorweek[129:158]
after_sept <- as.data.frame(after_sept)
```

# creating binary variable where 0 is when the date is before september 2015, and 1 when it is after
```{r}
month_or_week_clean$monthorweek <- ifelse(month_or_week_clean$monthorweek %in% after_sept, 0, 1)

```


# defining low and high earning income

```{r warning=FALSE}
earnings <- ourdata_norepeat %>%
  group_by(unitid, schname) %>%
  summarize(mean(as.numeric(md_earn_wne_p10.REPORTED.EARNINGS))) %>%
  na.omit(earnings, c("md_earn_wne_p10.REPORTED.EARNINGS"))
```
# finding mean and standard deviation of the data
```{r warning=FALSE}
mean_earnings <- mean(earnings$'mean(as.numeric(md_earn_wne_p10.REPORTED.EARNINGS))')
stdev_earnings <- sd(earnings$'mean(as.numeric(md_earn_wne_p10.REPORTED.EARNINGS))')

mean_earnings - stdev_earnings
mean_earnings + stdev_earnings
```
# creating binary variables for high and low earnings based off +-  1 standard deviation away from the mean
```{r}
colnames(earnings)[3] <- "avg_earnings"

earnings$high_income <- ifelse(earnings$'avg_earnings' >= 54141.91, 1, 0)
earnings$low_income <- ifelse(earnings$'avg_earnings' <= 30516.52, 1, 0)
```

#TEST 1
```{r}
ourdatafinal <- ourdata_norepeat %>%
  group_by(keynum) %>%
  summarise(standardized_index = (index - mean(index)) / sd(index), schname, standardized_index, CONTROL, keyword, monthorweek, keynum, opeid)
```


# TEST 2
```{r}
ourdatafinal <- ourdata_norepeat %>%
  group_by(keynum) %>%
  summarise(sd_index = (index - mean(index)) / sd(index), schname, CONTROL, 
            keyword, monthorweek, keynum, sd_index)
```


# TEST 3 standardizing the index variable for each term, subtract mean then divide by standard deviation 
```{r}
ourdatafinal <- ourdata %>%
  group_by(schname, keyword) %>%
  mutate(standardize_index = (index - mean(index)) / sd(index))

ourdatafinal <- ourdatafinal %>%
  relocate(standardize_index, .after = index)

```


# graphing

```{r}

```



